---
title: "SWORK 881 week02"
author: "Talin Gharibian "
date: "02/02/2022"
output: html_document
---


# load libraries
```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
```


# Data
```{r}
# load data
data.01 <- read_csv("Data/LHS_exercise_01272022.csv")

#how many observations and variables? 
```
#how many observations and variables? 
268 Observations; 8 Variables


```{r}
# create a new variable called `depression` by summing all ptsd variables. Use tidy approach
data.02 <- data.01%>%
  mutate(depression_score = ptsd1_1 + ptsd2_1 + ptsd3_1 + ptsd4_1)

```

```{r}
data.03 <- data.01%>%
  select(ptsd1_1, ptsd2_1, ptsd3_1, ptsd4_1, age, education, sex)%>%
  mutate(depression_score = rowSums(across(ptsd1_1:ptsd4_1)), na.rm=T)
```


```{r}
#run simple descriptive statistics for all variables
summary(data.03)
```

```{r}
mean(data.01$age)
sd(data.01$age)

ftable(data.01$sex)

ftable(data.01$education)

ftable(data.01$ptsd1_1)

ftable(data.01$ptsd2_1)

ftable(data.01$ptsd3_1)

ftable(data.01$ptsd4_1)
```


DESCRIBE THE DATA BRIEFLY HERE
Age mean (sd) : 32.11 (12.478)

Sex 
female : 164 (61.2%)
male : 104 (38.8%)
    
Education
1 - Less than high school : 18 (6.7%)
2 - Graduated high school : 91 (33.9%)
3 - Some college : 86 (32.1%)
4 - Graduated college or higher : 73 (27.2%)

PTSD 1
1 - Extremely : 121 (45.1%)
2 - Quite a bit : 72 (26.8%)
3 - Moderately : 34 (12.6%)
4 - A little bit : 24 (8.9%)
5 - Not at all : 17 (6.3%)

PTSD 2
1 : 89 (33.2%)
2 : 65 (24.2%)
3 : 47 (17.5%)
4 : 48 (17.9%)
5 : 19 (7.08%)

PTSD 3
1 : 122 (45.5%)
2 : 48 (17.9%)
3 : 35 (13.0%)
4 : 39 (14.5%)
5 : 24 (8.9%)

PTSD 4
1 : 119 (44.4%)
2 : 52 (19.4%)
3 : 30 (11.2%)
4 : 37 (13.8%)
5 : 30 (11.2%)


# Model and Visualization
> What is the effect of `age`, `sex`, and `education` on `depression score`? 

Create a linear regression model that answers the question above. Print the results of the output and model diagnostics.

```{r}
# lm
fit.1 <- lm(depression_score ~ age + education + factor(sex),
            data=data.03)
summary(fit.1)

```


```{r}
# model diagnostics
par(mfrow=c(2,2))
plot(fit.)

```


```{r}
# use ggplot to plot histogram of residuals (use 20 bins)
ggplot(fit.1,
       aes(fit.1$residuals))+
  geom_histogram(bins = 20)

```

# Discussion of Results

Describe and interpret your results, including any statistical information you believe important: 

We fit a linear regression model to measure the effect of `age`, `sex`, and `education` on `depression score.' Sex was not a significant predictor of depression score. Education was also not a significant predictor of depression score. Age was a statistically significant predictor of depression score (p value = 0.01). The residuals do not follow a normal distribution.
